Optimal design of experiment

### 2025-05-20 Article_Design_of_Experiment-Chamomile.tex
20.05

\usepackage{etoolbox}
\renewcommand\nomgroup[1]{%
	\item[\bfseries
	\ifstrequal{#1}{A}{Latin symbols}{%
		\ifstrequal{#1}{B}{Greek symbols}{%
			\ifstrequal{#1}{C}{Abberivations}{
	}}}%
	]}

\usepackage{csvsimple}

@@

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study investigates the process of chamomile oil extraction from flowers. A parameter-distributed model consisting of a set of partial differential equations is used to describe the governing mass transfer phenomena in a solid-fluid environment under supercritical conditions using carbon dioxide as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent are estimated from the Peng-Robinson equation of state. The empirical correlations used in the model are based on the set of laboratory experiments performed under multiple constant operating conditions: $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. A model-based design of experiments with D-optimality criterion is applied to improve the precision of the correlation parameters by designing a new dynamic experiment. The mass flow rate and inlet temperature are used as decision variables to maximize the Fisher Information embedded in the yield curve with respect to the empirical correlations.

@@

% ===================================================
% Bibliography
% ===================================================
%% Loading bibliography style file
\clearpage
\newpage
%\bibliographystyle{model1-num-names}
\bibliographystyle{unsrtnat}
\bibliography{mybibfile}

@@

\nomenclature[A]{\(P\)}{Pressure}
\nomenclature[A]{\(T\)}{Temperature}
\nomenclature[A]{\(T^{in}\)}{Inlet temperature}
\nomenclature[A]{\(T^{out}\)}{Outlet temperature}
\nomenclature[A]{\(F\)}{Mass flow rate}
\nomenclature[A]{\(u\)}{Superfical velocity}
\nomenclature[A]{\(v\)}{Linear velocity}
\nomenclature[A]{\(z\)}{Spatial direction}
\nomenclature[A]{\(e\)}{Internal energy}
\nomenclature[A]{\(e^y\)}{Output error}
\nomenclature[A]{\(h\)}{Enthalpy}
\nomenclature[A]{\(t\)}{Time}
\nomenclature[A]{\(t_f\)}{Total extraction time}
\nomenclature[A]{\(t_0\)}{Inital extraction time}
\nomenclature[A]{\(A\)}{Total cross section of the bed}
\nomenclature[A]{\(A_f\)}{Cross section of the bed accupied by the fluid}
\nomenclature[A]{\(c_f\)}{Concentration of solute in fluid phase}
\nomenclature[A]{\(c_{f0}\)}{Inital concentration of solute in fluid phase}
\nomenclature[A]{\(c_f^*\)}{Concentration of solute at the solid-fluid interface}
\nomenclature[A]{\(c_s\)}{Concentration of solute in solid phase}
\nomenclature[A]{\(c_{s0}\)}{Inital concentration of solute in solid phase}
\nomenclature[A]{\(c_s^*\)}{Concentration of solute at the solid-fluid interface}
\nomenclature[A]{\(c_p\)}{Concentration of solute in the core of a pore}
\nomenclature[A]{\(c_{pf}\)}{Concentration of solute in the pore opening}
\nomenclature[A]{\(D_e^M\)}{Axial diffusion coefficeint}
\nomenclature[A]{\(r_e\)}{Mass transfer kinetic term}
\nomenclature[A]{\(L\)}{Length of fixed bed}
\nomenclature[A]{\(j\)}{Objective function}
\nomenclature[A]{\(j_{ml}\)}{Maximum Likelihood}
\nomenclature[A]{\(S\)}{Sensitivity equations}
\nomenclature[A]{\(Q\)}{Weighting matrix}
\nomenclature[A]{\(R\)}{Control cost matrix}
%\nomenclature[A]{\(\textbf{G}\)}{Augumented system}
%\nomenclature[A]{\(\dot{S}\)}{Time derivative of sensitivity equations}
%\nomenclature[A]{\(\bar{S}\)}{Sensitivity matrix}
%\nomenclature[A]{\(\bar{J}_x\)}{Jacobian matrix with respect to the state space}
%\nomenclature[A]{\(\bar{J}_\Theta\)}{Jacobian matrix with respect to the paramters}
%\nomenclature[A]{\(\mathcal{L}\)}{Log-likelihood function}
\nomenclature[A]{\(l\)}{Characterisitic diemsinon of particles}
\nomenclature[A]{\(D_i\)}{Interanl diffusion coefficient}
\nomenclature[A]{\(D_i^R\)}{Reference value of interanl diffusion coefficient}
\nomenclature[A]{\(r\)}{Particle radius}
\nomenclature[A]{\(k_p\)}{Volumetric partition coefficient}
\nomenclature[A]{\(k_m\)}{Mass partition coefficient}
\nomenclature[A]{\(y\)}{Extraction yield}
\nomenclature[A]{\(Y\)}{Yield measurment}
\nomenclature[A]{\(x\)}{State vector}
\nomenclature[A]{\(G\)}{Vector of discretized differential equations}
\nomenclature[A]{\(p\)}{Probaility disribution model}
\nomenclature[A]{\(R_e\)}{Reynolds number}
\nomenclature[A]{\(\mathcal{F}\)}{Fisher information}

\nomenclature[B]{\(\rho_f\)}{Fluid density}
\nomenclature[B]{\(\Phi\)}{Bed porosity}
\nomenclature[B]{\(\rho_s\)}{Bulk density of solid}
\nomenclature[B]{\(\mu\)}{Sphericity coefficient}
\nomenclature[B]{\(\pi\)}{Probability desnsity}
\nomenclature[B]{\(\gamma\)}{Decaying function}
\nomenclature[B]{\(\Upsilon\)}{Decay coefficient}
\nomenclature[B]{\(\Theta\)}{Paramter space}
\nomenclature[B]{\(\theta\)}{vector of unknown parameters}
\nomenclature[B]{\(\epsilon\)}{Unobervable error}
\nomenclature[B]{\(\sigma\)}{Standard deviation}
\nomenclature[B]{\(\Sigma\)}{Covariance matrix}
\nomenclature[B]{\(\xi\)}{Vector of experimental condtions}
\nomenclature[B]{\(\Xi\)}{Matrix of experimental condtions}

>>>>

\usepackage{etoolbox}
\renewcommand\nomgroup[1]{%
	\item[\bfseries
	\ifstrequal{#1}{A}{}{%
		\ifstrequal{#1}{B}{Greek symbols}{%
			\ifstrequal{#1}{C}{Abberivations}{
	}}}%
	]}
\newcommand{\nomunit}[1]{%
	\renewcommand{\nomentryend}{\hspace*{\fill}#1}}

@@

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study investigates the process of chamomile oil extraction from flowers. A parameter-distributed model consisting of a set of partial differential equations is used to describe the governing mass transfer phenomena in a{ \color{blue}cylindrical packed bed with solid chamomile particles} under supercritical conditions using carbon dioxide as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent are estimated from the Peng-Robinson equation of state. The empirical correlations used in the model are based on the set of laboratory experiments performed under multiple constant operating conditions: ${\color{blue}31} - 40~^\circ$C, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. A model-based design of experiments with D-optimality criterion is applied to improve the precision of the correlation parameters by designing a new dynamic experiment. The mass flow rate and inlet temperature are used as decision variables to maximize the Fisher Information embedded in the yield curve with respect to the empirical correlations.

@@

% ===================================================
% Bibliography
% ===================================================
%% Loading bibliography style file
%\clearpage
%\newpage
%\bibliographystyle{model1-num-names}
\bibliographystyle{unsrtnat}
\bibliography{mybibfile}

@@

\nomenclature[A]{\(P\)}{Pressure \nomunit{bar}}
\nomenclature[A]{\(T\)}{Temperature \nomunit{K}}
\nomenclature[A]{\(T_{in}\)}{Inlet temperature \nomunit{K}}
\nomenclature[A]{\(T_{out}\)}{Outlet temperature \nomunit{K}}
\nomenclature[A]{\(F\)}{Mass flow rate \nomunit{kg/s}}
\nomenclature[A]{\(u\)}{Superficial velocity \nomunit{m/s}}
\nomenclature[A]{\(v\)}{Linear velocity \nomunit{m/s}}
\nomenclature[A]{\(z\)}{Spatial direction \nomunit{m}}
\nomenclature[A]{\(e\)}{Internal energy \nomunit{J/kg}}
%\nomenclature[A]{\(e^y\)}{Output error}
\nomenclature[A]{\(h\)}{Enthalpy \nomunit{kJ/kg}}
\nomenclature[A]{\(t\)}{Time \nomunit{s}}
\nomenclature[A]{\(t_f\)}{Total extraction time \nomunit{s}}
\nomenclature[A]{\(t_0\)}{Initial extraction time \nomunit{s}}
\nomenclature[A]{\(A\)}{Total cross-section of the bed \nomunit{m$^2$}}
\nomenclature[A]{\(A_f\)}{Cross-section of the bed occupied by the fluid \nomunit{m$^2$}}
\nomenclature[A]{\(c_f\)}{Concentration of solute in the fluid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_{f0}\)}{Initial concentration of solute in the fluid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_f^*\)}{Concentration of solute at the solid-fluid interface \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_s\)}{Concentration of solute in the solid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_{s0}\)}{Initial concentration of solute in the solid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_s^*\)}{Concentration of solute at the solid-fluid interface \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_p\)}{Concentration of solute in the core of a pore \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_{pf}\)}{Concentration of solute at the pore opening \nomunit{kg/m$^3$}}
\nomenclature[A]{\(D_e^M\)}{Axial diffusion coefficient \nomunit{m$^2$/s}}
\nomenclature[A]{\(r_e\)}{Mass transfer kinetic term \nomunit{kg/m$^3$/s}}
\nomenclature[A]{\(L\)}{Length of the fixed bed \nomunit{m}}
\nomenclature[A]{\(j\)}{Objective function \nomunit{-}}
\nomenclature[A]{\(j_{ml}\)}{Maximum Likelihood \nomunit{-}}
\nomenclature[A]{\(S\)}{Sensitivity equations \nomunit{-}}
\nomenclature[A]{\(Q\)}{Weighting matrix \nomunit{-}}
\nomenclature[A]{\(R\)}{Control cost matrix \nomunit{-}}
%\nomenclature[A]{\(\textbf{G}\)}{Augumented system}
%\nomenclature[A]{\(\dot{S}\)}{Time derivative of sensitivity equations}
%\nomenclature[A]{\(\bar{S}\)}{Sensitivity matrix}
%\nomenclature[A]{\(\bar{J}_x\)}{Jacobian matrix with respect to the state space}
%\nomenclature[A]{\(\bar{J}_\Theta\)}{Jacobian matrix with respect to the paramters}
%\nomenclature[A]{\(\mathcal{L}\)}{Log-likelihood function}
\nomenclature[A]{\(l\)}{Characteristic dimension of particles \nomunit{m}}
\nomenclature[A]{\(D_i\)}{Internal diffusion coefficient \nomunit{m$^2$/s}}
\nomenclature[A]{\(D_i^R\)}{Reference value of internal diffusion coefficient \nomunit{m$^2$/s}}
\nomenclature[A]{\(r\)}{Particle radius \nomunit{m}}
\nomenclature[A]{\(k_p\)}{Volumetric partition coefficient \nomunit{-}}
\nomenclature[A]{\(k_m\)}{Mass partition coefficient \nomunit{-}}
\nomenclature[A]{\(y\)}{Extraction yield \nomunit{g}}
%\nomenclature[A]{\(Y\)}{Yield measurement \nomunit{g}}
\nomenclature[A]{\(x\)}{State vector}
\nomenclature[A]{\(G\)}{Vector of discretized differential equations}
\nomenclature[A]{\(p\)}{Probaility disribution model}
\nomenclature[A]{\(R_e\)}{Reynolds number}
\nomenclature[A]{\(\mathcal{F}\)}{Fisher information}

\nomenclature[B]{\(\rho_f\)}{Fluid density\nomunit{kg/m$^3$}}
\nomenclature[B]{\(\Phi\)}{Bed porosity \nomunit{-}}
\nomenclature[B]{\(\rho_s\)}{Bulk density of solid\nomunit{kg/m$^3$}}
\nomenclature[B]{\(\mu\)}{Sphericity coefficient \nomunit{-}}
\nomenclature[B]{\(\pi\)}{Probability desnsity \nomunit{-}}
\nomenclature[B]{\(\gamma\)}{Decaying function \nomunit{-}}
\nomenclature[B]{\(\Upsilon\)}{Decay coefficient \nomunit{-}}
\nomenclature[B]{\(\Theta\)}{Paramter space \nomunit{-}}
\nomenclature[B]{\(\theta\)}{vector of unknown parameters \nomunit{-}}
\nomenclature[B]{\(\epsilon\)}{Unobervable error \nomunit{-}}
\nomenclature[B]{\(\sigma\)}{Standard deviation \nomunit{-}}
\nomenclature[B]{\(\Sigma\)}{Covariance matrix \nomunit{-}}
\nomenclature[B]{\(\xi\)}{Vector of experimental condtions \nomunit{-}}
\nomenclature[B]{\(\Xi\)}{Matrix of experimental condtions \nomunit{-}}

### 2025-05-20 Sections/DOE.tex
20.05

	Building on the work of\citet{Walter2010} and \citet{Himmelblau1970}, Bayes' Theorem can be employed to define a general cost function:

@@

	The aim of experimental design is to maximize the posterior probability of estimating $\theta$ ($p\left(\theta \mid Y, \Xi \right)$) given the prior distribution of $\theta$ ($p(\theta)$) obtained from previous experiments, and the likelihood of obtaining observations $Y$ given $\theta$ and experimental conditions $\Xi$. The new observations $Y$, under a set of experimental conditions $\Xi$, are assumed to be normally distributed around the model output $y$, with measurement error covariance $\Sigma_Y$. Similarly, $\theta$ is assumed to be normally distributed around the true parameter values $\hat{\theta}$, with covariance matrix $\Sigma_\theta$ estimated from previous experiments. The probabilities can be defined as

	{\footnotesize
		\begin{align} 
			p\left(Y \mid \theta, \Xi \right) &= \frac{|\Sigma_Y|^{-1/2}}{\left(2\pi\right)^{n_Y/2}} \exp\left( -\frac{1}{2} \left(Y - y\right)^\top \Sigma_Y^{-1} \left(Y - y\right) \right), \\ 
			p\left(\theta\right) &= \frac{|\Sigma_\theta|^{-1/2}}{\left(2\pi\right)^{n_\theta/2}} \exp\left( -\frac{1}{2} \left(\theta - \hat{\theta}\right)^\top \Sigma_\theta^{-1} \left(\theta - \hat{\theta}\right) \right),
		\end{align}
	}

@@

	By linearizing the model output around $\hat{\theta}$, the model output becomes: $y = y(\hat{\theta}) + J \Delta_\theta$, where $J = \frac{\partial y(\hat{\theta})}{\partial \theta}$ is the Jacobian matrix evaluated at $\hat{\theta}$, and $\Delta_\theta = \theta - \hat{\theta}$. To simplify the equation, we introduce the residual term $\Delta_y = Y - y(\hat{\theta})$. Substituting these into the expression for $\ln p\left(\theta \mid Y, \Xi \right)$:

@@

	The Fisher Information $\mathcal{F}$ measures the amount of information that an observable random variable carries about an unknown parameter of the distribution that models the random variable. It is related to the negative expected value of the second derivative (or curvature) of the log-likelihood function with respect to the parameter, providing a measure of how "sensitive" the likelihood is to changes in the parameter value.

@@

	Since the likelihood is Gaussian and linearized, the second derivative of the log-likelihood with respect to $\theta$ simplifies to $J^\top \Sigma_Y^{-1} J$, and its negative expected value yields the same expression because the expectation of the Hessian is constant in this linear approximation.
	\begin{comment}

	Under the hypotheses that:

	\begin{itemize}
		\item The set of all data vectors $Y$ with $p(Y \mid \theta, \Xi) > 0$ does not depend on $\theta$.
		\item The partial derivatives $\frac{\partial p(Y \mid \theta, \Xi)}{\partial \theta_i}$ for $i=1,2,...,n_\theta$ are absolutely integrable.
		\item The expected value $\mathop{\mathbb{E}}_{Y \mid \theta, \Xi} \left[ \frac{\partial^2 \ln p \left( Y \mid \theta, \Xi \right)}{\partial \theta \partial \theta^\top} \right]$ exists and is invertible.
	\end{itemize}

	\end{comment}
	The Cramér-Rao inequality provides that the covariance of any unbiased estimator $\hat{\theta}$ satisfies:

@@

	In other words, the precision with which we can estimate $\theta$ is limited by the Fisher Information of the likelihood function. Based on the Cramér-Rao inequality, the Fisher Information matrix can be used to calculate the covariance matrices associated with maximum-likelihood estimates.

@@

	A general class of design of experiments (DOE) optimality criteria is given by:

	{\footnotesize 
		\begin{align} 
			\kappa_k\left( A^{-1}(\theta) \right) &= \left[ \frac{1}{n_\theta} \operatorname{trace}\left( Q A^{-1}(\theta, \Xi) Q^\top \right)^k \right]^{1/k} & \text{if} \det A^{-1} \neq 0, \nonumber \\ 
			\kappa_k\left( A^{-1}(\theta) \right) &= \infty & \text{if} \det A^{-1} = 0
		\end{align} }

@@

	The D-optimality criterion originates from the geometric interpretation of the determinant. A two-dimensional matrix with real number entries can represent two linear maps: one mapping the standard basis vectors to the rows and the other to the columns. In both cases, the images of the basis vectors form a parallelogram that represents the image of the unit square under the mapping. The absolute value of the determinant is the area of this parallelogram, reflecting the scale factor by which the matrix transforms areas. The signed value of the determinant indicates the oriented area of the parallelogram, which is negative when the angle from the first to the second defining vector is clockwise. In the multi-dimensional case, the matrix maps the n-cube to an n-dimensional parallelotope. The determinant provides the signed n-dimensional volume of this parallelotope, describing the volume scaling factor of the linear transformation produced by the matrix. Based on this geometric interpretation, a D-optimal experiment minimizes the volume of the asymptotic confidence ellipsoids for the parameters. Figure \ref{fig:score_fun} illustrates the graphical representation of different optimality conditions.

	\begin{figure}[!h]
		\centering
		\resizebox{0.4\textwidth}{!}{%
			\begin{tikzpicture}
				% Ellipse parameters
				\def\ellipsecenter{(3,3)}
				\def\ellipseA{(0.75,5.25)} % End point for A-optimality arrow
				\def\ellipseB{(5.13,5.13)} % End point for E-optimality arrow
				\def\ellipseC{(4,1.2)}      % End point for D-optimality arrow

@@

	Details on the process model can be found in [article 1]. The model's empirical correlations are derived based on laboratory experiments conducted under various but constant operating conditions: $30^\circ\text{C} - 40^\circ\text{C}$, $100 - 200$ bar, and $3.33 \cdot 10^{-5} - 6.67 \cdot 10^{-5}$ kg/s. This study employs a model-based design of experiments technique to design an experiment with dynamically changing conditions to improve the precision of the correlation for $D_i$. The decision variables are adjusted every 10 minutes and are kept constant within each interval (piecewise constant controls). These controls have lower and upper bounds that match the validated range of the process model detailed in [article 1]. The sampling time is set to 5 minutes, and the total extraction time is 600 minutes. The standard deviation $\sigma^2$ was estimated to be 0.10.

	The analyzed correlations consist of two independent variables: Reynolds number and mass flow rate. The Reynolds number is a function of fluid thermodynamic properties and velocity, and the fluid density can be affected by temperature or pressure changes. In this work, the decision variables are the inlet temperature ($T^{\text{in}}$) and mass flow rate ($F$), while the pressure is assumed to remain constant. The initial state is considered isothermal, so $T^0 = T^{\text{in}}(t=0)$. The optimal design of experiment problem is solved for multiple values of pressure, namely 100, 125, 150, 175, and 200 bar.

	In a real system, the mass flow rate and inlet temperature cannot be changed instantaneously, as they depend on the dynamics of the pump and heat exchanger, respectively. To prevent bang-bang-like control between the lower and upper bounds of the decision variables, a penalty term is introduced. Similar to many control problems, such as the Linear Quadratic Regulator, a quadratic penalty term is added to the cost function. This penalty increases the cost function value when there are rapid changes in the decision variables. The matrix $R$ represents the control cost matrix, with its entries chosen so that the control costs are one order of magnitude lower than $-\ln j_D$. The problem formulation is given by Equation \ref{EQ:Formulation_2}.

	{\footnotesize 
		\begin{equation} 
			\begin{aligned} \Xi^* &= \arg \min_{ T^{\text{in}}(t), F(t) } \int_{t_0}^{t_f} \left[ - \ln j_D(\Xi, \dot{x}) + u(t)^\top R u(t) \right] dt \\ 
				\text{subject to} \quad & \dot{x} = G(x, t, \theta; \Xi), \\ 
				& t_0 = 0 \text{min} \\ 
				& t_f = 600 \text{min} \\ 
				& T^{0} = T^{\text{in}}(t=0), \\ 
				& P(t) \in \left\{100, 125, 150, 175, 200\right\} \text{bar}, \\ 
				%& R = \begin{bmatrix} 10^{-1} & 0 \ 0 & 10^{-2} \end{bmatrix}, \\ 
				%& u(t) = \begin{bmatrix} \dfrac{dF(t)}{dt} \ \dfrac{dT^{\text{in}}(t)}{dt} \end{bmatrix}, \\
				& 30^\circ\text{C} \leq T^{\text{in}}(t) \leq 40^\circ\text{C}, \\ 
				& 3.33 \cdot 10^{-5} \leq F(t) \leq 6.67 \cdot 10^{-5} 
				\end{aligned} 
				\label{EQ:Formulation_2} 
			\end{equation} }

>>>>

	Building on the work of \citet{Walter2010} and \citet{Himmelblau1970}, the Bayes' Theorem can be employed to define a general cost function:

@@

	The aim of experimental design is to maximize the posterior probability of estimating $\theta$ ($p\left(\theta \mid Y, \Xi \right)$) given the prior distribution of $\theta$ ($p(\theta)$) obtained from previous experiments, and the likelihood of obtaining observations $Y~\left(p\left(Y \mid \theta, \Xi\right)\right)$ given $\theta$ and experimental conditions $\Xi$. The new observations $Y$, under a set of experimental conditions $\Xi$, are assumed to be normally distributed around the model output $y$, with measurement error covariance $\Sigma_Y$. Similarly, $\theta$ is assumed to be normally distributed around the true parameter values $\hat{\theta}$, with covariance matrix $\Sigma_\theta$ estimated from previous experiments. These probabilities are:

	{\footnotesize
		\begin{align} 
			p\left(Y \mid \theta, \Xi \right) &= \frac{|\Sigma_Y|^{-1/2}}{\left(2\pi\right)^{n_Y/2}} \exp\left( -\frac{1}{2} \left(Y - y\right)^\top \Sigma_Y^{-1} \left(Y - y\right) \right) \\
			p\left(\theta\right) &= \frac{|\Sigma_\theta|^{-1/2}}{\left(2\pi\right)^{n_\theta/2}} \exp\left( -\frac{1}{2} \left(\theta - \hat{\theta}\right)^\top \Sigma_\theta^{-1} \left(\theta - \hat{\theta}\right) \right)
		\end{align}
	}

@@

	By linearizing the model output around $\hat{\theta}$, the model output becomes: $y = y(\hat{\theta}) + J \Delta_\theta$, where $J = \frac{\partial y(\hat{\theta})}{\partial \theta}$ is the Jacobian matrix evaluated at $\hat{\theta}$, and $\Delta_\theta = \theta - \hat{\theta}$. To simplify the equation, the residual term $\Delta_y = Y - y(\hat{\theta})$ is introduced. Substituting these into the expression for $\ln p\left(\theta \mid Y, \Xi \right)$:

@@

	The Fisher Information $\mathcal{F}$ measures the amount of information that an observable random variable carries about an unknown parameter of the distribution that models the random variable. It is related to the negative expected value of the second derivative of the log-likelihood function with respect to the parameter, providing a measure of how "sensitive" the likelihood is to changes in the parameter value.

@@

	Since the likelihood is Gaussian and linearized, the second derivative of the log-likelihood with respect to $\theta$ simplifies to $J^\top \Sigma_Y^{-1} J$, and its negative expected value yields the same expression because the expectation of the Hessian is constant in this linear approximation.

	The Cramér-Rao inequality provides that the covariance of any unbiased estimator $\hat{\theta}$ satisfies:

@@

	The precision with which a parameter $\theta$ can be estimated is limited by the Fisher Information of the likelihood function. Based on the Cramér-Rao inequality, the Fisher Information matrix can be used to calculate the covariance matrices associated with maximum-likelihood estimates.

@@

	\begin{figure}[!h]
		\centering
		\resizebox{0.9\columnwidth}{!}{%
			\begin{tikzpicture}
				% Ellipse parameters
				\def\ellipsecenter{(3,3)}
				\def\ellipseA{(0.75,5.25)} % End point for A-optimality arrow
				\def\ellipseB{(5.13,5.13)} % End point for E-optimality arrow
				\def\ellipseC{(4,1.2)}      % End point for D-optimality arrow

	A general class of design of experiments (DOE) optimality criteria is given by Equation \ref{EQ:DOE_general}. Some of the DOE criterion have graphical interpretation as shown in Figure \ref{fig:score_fun}.

	{\footnotesize 
		\begin{align} \label{EQ:DOE_general}
			\kappa_k\left( A^{-1}(\theta) \right) &= \left[ \frac{1}{n_\theta} \operatorname{trace}\left( Q A^{-1}(\theta, \Xi) Q^\top \right)^k \right]^{1/k} & \text{if} \det A^{-1} \neq 0, \nonumber \\ 
			\kappa_k\left( A^{-1}(\theta) \right) &= \infty & \text{if} \det A^{-1} = 0
		\end{align} }

@@

	The D-optimality (D-OED) criterion originates from the geometric interpretation of the determinant. A two-dimensional matrix with real number entries can represent two linear maps: one mapping the standard basis vectors to the rows and the other to the columns. In both cases, the images of the basis vectors form a parallelogram that represents the image of the unit square under the mapping. The absolute value of the determinant is the area of this parallelogram, reflecting the scale factor by which the matrix transforms areas. The signed value of the determinant indicates the area of the parallelogram, which is negative when the angle from the first to the second defining vector is clockwise. In the multi-dimensional case, the matrix maps the n-cube to an n-dimensional parallelotope. The determinant provides the signed n-dimensional volume of this parallelotope, describing the volume scaling factor of the linear transformation produced by the matrix. Based on this geometrical interpretation, a D-optimal experiment minimizes the volume of the asymptotic confidence ellipsoids for the parameters.

@@

	Details on the process model can be found in \citet{Sliczniuk2024}. The model's empirical correlations are derived based on laboratory experiments conducted under various but constant operating conditions: $31 - 40~^\circ\text{C}$, $100 - 200$ bar, and $3.33 \cdot 10^{-5} - 6.67 \cdot 10^{-5}$ kg/s. This study employs the process model obtained previously to design a set of supplementary experiments with dynamically changing conditions to improve the precision of the correlation for $D_i$. The analyzed correlations consist of three independent variables: mass flow rate, pressure and inlet temperature.

	In this work, the decision variables are the inlet temperature ($T^{\text{in}}$) and mass flow rate ($F$), while the pressure is assumed to remain constant during a batch. The optimal design of experiment problem is solved for multiple values of pressure: 100, 125, 150, 175, and 200 bar. The initial state is considered isothermal, so $T^0(z) = T^{\text{in}}(t=0)$.

	The decision variables are adjusted every 10 minutes and are kept constant within each interval (piecewise constant controls). These controls have lower and upper bounds that match the validated range of the process model detailed by \citet{Sliczniuk2024}. The sampling time is assumed to be 5 minutes, and the total extraction time is 600 minutes. The standard deviation $\sigma^2$ was estimated to be 0.03.

	In a real system, the mass flow rate and inlet temperature cannot be changed instantaneously to any arbitrary value, as they depend on the dynamics of the pump and heat exchanger, respectively. To prevent bang-bang type of control, a penalty term is introduced. Inspired by control problems, such as the Linear Quadratic Regulator, a quadratic penalty term is added to the cost function. This penalty increases the cost function value when there are rapid changes in the decision variables. The matrix $\bar{R}$ represents the control cost matrix, with its entries chosen so that the control costs are one order of magnitude lower than $-\ln j_D$. The problem formulation is given by Equation \ref{EQ:Formulation_2}.

	{\footnotesize 
		\begin{equation}
			\begin{aligned}
				\Xi^* &= \arg \min_{T^{\text{in}}(t), F(t)} 
				\int_{t_0}^{t_f} \left[ - \ln j_D(\Xi, \dot{x}) + 
				\frac{du(t)^\top}{dt} \, \bar{R} \, \frac{du(t)}{dt} \right] dt, \\
				\text{subject to:} \quad 
				& \dot{x} = G(x, t, \theta; \Xi), \\
				& t_0 = 0 \, \text{min}, \quad t_f = 600 \, \text{min}, \\
				& T^0 = T^{\text{in}}(t=0), \\
				& P(t) \in \{100, 125, 150, 175, 200\} \, \text{bar}, \\
				& 31^\circ\text{C} \leq T^{\text{in}}(t) \leq 40^\circ\text{C}, \\
				& 3.33 \cdot 10^{-5} \leq F(t) \leq 6.67 \cdot 10^{-5}, \\
				& \bar{R} = 
				\begin{bmatrix} 
					0.1 & 0 \\ 
					0 & 0.01 
				\end{bmatrix}, \\
				& \frac{du(t)}{dt} = 
				\begin{bmatrix} 
					\dfrac{dF(t)}{dt} \\ 
					\dfrac{dT^{\text{in}}(t)}{dt} 
				\end{bmatrix}.
			\end{aligned}
			\label{EQ:Formulation_2}
			\end{equation} }

### 2025-05-20 Sections/Literature_Review.tex
20.05

	\citet{Ramandi2011} conducted a study to identify the variables with the greatest influence on the extraction yield of fatty acids from \textit{Borago officinalis L.} flowers using Supercritical Fluid Extraction (SFE). A Full Factorial Design was employed as a screening method, resulting in a design matrix of 32 runs ($2^5$) carried out randomly to minimize the impact of extraneous or nuisance variables. The low and high values for the factors were selected based on previous research. After determining the most significant variables, a Central Composite Design was applied to three factors—temperature, pressure, and modifier volume—to optimize the SFE conditions. This allowed for the modeling of the response surface by fitting a second-order polynomial.

@@

	After an initial dataset has been collected and fitted to a mathematical model, the model undergoes further analysis. Additional experiments may be designed and conducted to differentiate between competing models that passed the preliminary tests. Once inadequate models are rejected, the remaining model may undergo another round of experiment design to enhance the precision of its parameters. This paper focuses on the final step of the validation procedure, known as model-based Design of Experiments (m-DoE), aimed at improving parameter precision. To the authors' knowledge, the m-DoE has not been applied to any case of supercritical extraction, so the further literature review provides examples of application in crystallization and pharmacology processes.

>>>>

	\citet{Ramandi2011} conducted a study to identify the variables with the greatest influence on the extraction yield of fatty acids from \textit{Borago officinalis L.} flowers using Supercritical Fluid Extraction (SFE). A Full Factorial Design was employed as a screening method, resulting in a design matrix of 32 runs ($2^5$) carried out randomly to minimize the impact of external factors. The low and high values for the factors were selected based on previous research. After determining the most significant variables, a Central Composite Design was applied to three factors—temperature, pressure, and volume of methanol co-solvent to optimize the SFE conditions. This allowed for the modeling of the response surface by fitting a second-order polynomial.

@@

	After an initial dataset has been collected and fitted to a mathematical model, the model undergoes further analysis. Additional experiments may be needed to differentiate between competing models that passed the preliminary tests. Once inadequate models are rejected, the remaining model may undergo another round of experiment design to enhance the precision of its parameters. This paper focuses on the final step of the validation procedure, known as model-based Design of Experiments (m-DoE), aimed at improving parameter precision. To the authors' knowledge, the m-DoE has not been applied to any case of supercritical extraction, so the further literature review provides examples of application in crystallization and pharmacology processes.
